GSEA datasets for FDR benchmarking

### 2017-10-27 datasets/GSEA/gsea.Rmd
Added functions to plot rejections vs alpha. Updated GSEA analysis.

```{r deseq2}

@@

```{r, fig.height=3.5, fig.width=4, message=FALSE}
library(ggplot2)
library(magrittr)
library(cowplot)

ggplot( goRes,
       aes( log10(numInCat), -log10(over_represented_pvalue) ) ) +
    geom_hex(aes( alpha=log2(..count..)), fill="black", color="black", bins=100, show.legend=FALSE) +
    ylim(0, 12) + ggtitle("Over-represented gene sets") +
    xlab(expression(log[10]~"(# of genes)")) +
    ylab(expression(-log[10]~"(p-value)"))

@@

```{r, message=FALSE}
library(IHW)
goRes$groups <- groups_by_filter( goRes$numInCat, 10 )

@@

dplyr:::filter( goRes, numInCat > 10, numDEInCat > 0 ) %>%
  dplyr:::mutate( groups = groups_by_filter( numInCat, 4 ) ) %>%
  ggplot(aes(x=over_represented_pvalue)) + 
  geom_histogram(binwidth = 0.025, boundary = 0) +
  facet_wrap( ~ groups, nrow = 2) +
  ggtitle("Distribution of p-values") +
  ylab("Frequency") + xlab("p-value")

@@

```{r}

library(R.utils)
sourceDirectory("../R/")

designBen <- initializeBenchDesign()
res <- dplyr:::select( goRes, c("over_represented_pvalue", "numInCat") ) %>%
  dplyr:::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
res$pval <- pmin( res$pval, 1 )

sGSEA <- designBen %>% buildBench( res )
assayNames(sGSEA) <- names( rowData(sGSEA) ) <-  names(sGSEA@performanceMetrics) <- "qvalue"

sGSEA <- addDefaultMetrics( sGSEA )

estimatePerformanceMetrics( sGSEA, alpha=seq(0, 0.6, 0.1), tidy=TRUE ) %>%
  dplyr:::filter(pkg_name == "IHW") %>%
  head

  dplyr:::select(blabel, key, value, assay, performanceMetric, alpha ) %>%
  dplyr:::filter(blabel!="unadjusted") %>%
  ggplot( aes(alpha, value, col=blabel) ) +
  geom_line() +
  geom_point()

>>>>

```{r deseq2, message=FALSE}

@@

```{r informativeCovariate, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
library(ggplot2)
library(magrittr)
library(cowplot)
library(R.utils)
library(scales)
sourceDirectory("../R/")

rank_scatter( dat=goRes, pval="over_represented_pvalue", 
              covariate="numInCat", bins=50, funx=log2, 
              funfill=log10_trans()) +
    ylim(0, 12) + ggtitle("Over-represented gene sets") +
    xlab( expression(log[10]~"(# of genes)")) +
    ylab( expression(-log[10]~"(p-value)") )

@@

```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}

library(dplyr)

@@

dplyr:::filter( goRes, numDEInCat > 0, numInCat > 15 ) %>%
  strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=7)

@@

```{r runningBenckmark, message=FALSE}
designBen <- initializeBenchDesign()
res <- dplyr:::select( goRes, c("over_represented_pvalue", "numInCat") ) %>%
  dplyr:::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
res$pval <- pmin( res$pval, 1 )
sGSEA <- designBen %>% buildBench( res )
assayNames(sGSEA) <- "qvalue"
sGSEA <- addDefaultMetrics( sGSEA )
```

```{r plottingResults}
rejections_scatter( sGSEA, as_fraction=FALSE )

stopifnot( all(assays(sGSEA)[["qvalue"]][,"unadjusted"] == pmin(goRes$over_represented_pvalue, 1) ) )
rowData(sGSEA)$numInCat <- goRes$numInCat

rejection_scatter_bins( sGSEA, "numInCat", threshold=15, bins=4, as_fraction=TRUE )
```
